---
title: "Project1"
author: "Vito Frank Leonardo"
date: "6/14/2021"
output: 
  html_document:
    toc: yes
    toc_float: no
    toc_depth: 3
---
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages Needed:
```{r}
library(httr)
library(jsonlite)
```


You should write functions to contact the NHL records API for the endpoints listed below. The
functions should return well-formatted, parsed data (usually a data frame). Where possible, the user should have the option to specify the franchise of choice by both name and ID number - you’ll need to map the names to ID numbers yourself

```{r cars}
Base_URL <- "https://records.nhl.com/site/api"

FranchiseInfo <- function(ID=NULL, Name=NULL){
  franchise <- "/franchise"
  req <- GET(paste0(Base_URL,franchise)) # from httr library
  data <- (content(req)) # from httr library

  df <- data.frame(matrix(rep(0,8), nrow=1))
  colnames(df) <- names(data$data[[2]])
  for( i in 1:length(data$data)){
    datalist <- data$data[[i]]
    temp <- data.frame(matrix(datalist, nrow=1))
    df[i,] <- temp
  }
  
  if(!is.null(ID) & !is.null(Name)){
    results <- filter(df, id %in% ID | teamCommonName %in% Name )
    if(dim(results)[1]!=0){
      return(results)
    }
  }

  if(!is.null(ID)){
    results <- filter(df, id %in% ID)
    if(dim(results)[1]!=0){
      return(results)
    } else if(is.null(Name)){
      stop("ID is not in this dataset. Choose an ID between 1 and 39")
    } else if(!is.null(Name)){
        results <- filter(df, teamCommonName %in% Name )
        if(dim(results)[1]!=0){
        return(results)
       } else stop("Neither ID nor Name are in the dataset. Try again")
    }
  }
    if(!is.null(Name)){
      results <- filter(df, teamCommonName %in% Name )
      if(dim(results)[1]!=0){
        return(results)
      } else stop("This name does not exist. Check spelling, captilize only the first letter, and use quotation marks")
    }
    
    return(df)
}
  
  
 

teamTotals <- "/franchise-team-totals"


FranchiseSeasonRecords <- paste0("/franchise-season-records?cayenneExp=franchiseId=",ID)
FranchiseGoalieRecords <- paste0("/franchise-goalie-records?cayenneExp=franchiseId=",ID)
FranchiseSkatorRecords <- paste0("/franchise-skater-records?cayenneExp=franchiseId=",ID)
FranchiseAdminHistory <- paste0("franchise-detail?cayenneExp=mostRecentTeamId=",ID)



```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
