---
title: "Project1"
author: "Vito Frank Leonardo"
date: "6/14/2021"
output: 
  html_document:
    toc: yes
    toc_float: no
    toc_depth: 3
---
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages Needed:
```{r}
library(httr)
library(jsonlite)
```


You should write functions to contact the NHL records API for the endpoints listed below. The
functions should return well-formatted, parsed data (usually a data frame). Where possible, the user should have the option to specify the franchise of choice by both name and ID number - you’ll need to map the names to ID numbers yourself

```{r BaseURL}
Base_URL <- "https://records.nhl.com/site/api"
```

```{r Franchise}
FranchiseInfo <- function(ID=NULL, Name=NULL){
  franchise <- "/franchise"
  req <- GET(paste0(Base_URL,franchise)) # from httr library
  data <- content(req) # from httr library

  df <- data.frame(matrix(rep(0,8), nrow=1))
  colnames(df) <- names(data$data[[2]])
  for( i in 1:length(data$data)){
    datalist <- data$data[[i]]
    temp <- data.frame(matrix(datalist, nrow=1))
    df[i,] <- temp
  }
  
  if(!is.null(ID) & !is.null(Name)){
    results <- filter(df, id %in% ID | teamCommonName %in% Name )
    if(dim(results)[1]!=0){
      return(results)
    }
  }

  if(!is.null(ID)){
    results <- filter(df, id %in% ID)
    if(dim(results)[1]!=0){
      return(results)
    } else if(is.null(Name)){
      stop("ID is not in this dataset. Choose an ID between 1 and 39")
    } else if(!is.null(Name)){
        results <- filter(df, teamCommonName %in% Name )
        if(dim(results)[1]!=0){
        return(results)
       } else stop("Neither ID nor Name are in the dataset. Try again")
    }
  }
    if(!is.null(Name)){
      results <- filter(df, teamCommonName %in% Name )
      if(dim(results)[1]!=0){
        return(results)
      } else stop("This name does not exist. Check spelling, captilize only the first letter, and use quotation marks")
    }
    
    return(df)
}
```

```{r teamTotals}

TeamTotals <- function(ID=NULL, Name=NULL){
  franchise <- "/franchise"
  req <- GET(paste0(Base_URL,franchise)) # from httr library
  data <- content(req) # from httr library

  df <- data.frame(matrix(rep(0,length(data$data[[2]])), nrow=1))
  colnames(df) <- names(data$data[[2]])
  for( i in 1:length(data$data)){
    datalist <- data$data[[i]]
    temp <- data.frame(matrix(datalist, nrow=1))
    df[i,] <- temp
  }
  
  NamesDf <- df %>% 
    select(id, fullName, teamCommonName, teamPlaceName, teamAbbrev) %>%
    rename(franchiseId = id)
  
  teamTotals <- "/franchise-team-totals"
  req <- GET(paste0(Base_URL,teamTotals)) # from httr library
  data <- (content(req)) # from httr library

  df <- data.frame(matrix(rep(0,length(data$data[[1]])), nrow=1))
  colnames(df) <- names(data$data[[2]])
  for( i in 1:length(data$data)){
    datalist <- data$data[[i]]
    temp <- data.frame(matrix(datalist, nrow=1))
    df[i,] <- temp
  }
  
  df <- df %>% left_join(NamesDf,"franchiseId")
  
  if(!is.null(ID) & !is.null(Name)){
    results <- filter(df, franchiseId %in% ID | teamCommonName %in% Name )
    if(dim(results)[1]!=0){
      return(results)
    }
  }

  if(!is.null(ID)){
    results <- filter(df, id %in% ID)
    if(dim(results)[1]!=0){
      return(results)
    } else if(is.null(Name)){
      stop("ID is not in this dataset. Choose an ID between 1 and 39")
    } else if(!is.null(Name)){
        results <- filter(df, teamCommonName %in% Name )
        if(dim(results)[1]!=0){
        return(results)
       } else stop("Neither ID nor Name are in the dataset. Try again")
    }
  }
    if(!is.null(Name)){
      results <- filter(df, teamCommonName %in% Name )
      if(dim(results)[1]!=0){
        return(results)
      } else stop("This name does not exist. Check spelling, captilize only the first letter, and use quotation marks")
    }
    
    return(df)
}

  
```
 
```{r Season}

SeasonRecords <- function(ID =NULL, Name = NULL){
  franchise <- "/franchise"
  req <- GET(paste0(Base_URL,franchise)) # from httr library
  data <- content(req) # from httr library

  df <- data.frame(matrix(rep(0,length(data$data[[2]])), nrow=1))
  colnames(df) <- names(data$data[[2]])
  for( i in 1:length(data$data)){
    datalist <- data$data[[i]]
    temp <- data.frame(matrix(datalist, nrow=1))
    df[i,] <- temp
  }

  NamesDf <- df %>% 
    select(id, teamCommonName) %>%
    filter(teamCommonName %in% Name)
  ID <- unique(union(ID,unlist(NamesDf$id )))
  
  
  if(is.null(ID)){
    ID <- c(1:39)
  }
  FranchiseSeasonRecords <- paste0("/franchise-season-records?cayenneExp=franchiseId=",ID[1])
  req <- GET(paste0(Base_URL,FranchiseSeasonRecords)) # from httr library
  data <- content(req) # from httr library
  df <- data.frame(data$data)
  for(i in 1:length(ID)){
    FranchiseSeasonRecords <- paste0("/franchise-season-records?cayenneExp=franchiseId=",ID[i])
    req <- GET(paste0(Base_URL,FranchiseSeasonRecords)) # from httr library
    data <- content(req) # from httr library
    temp <- data.frame(matrix(data$data[[1]],nrow = 1))
    colnames(temp) <- names(df)
    df <- rbind(df,temp)
  }
  return(df)
  } 
```

```{r}

GoalieRecords <- function(ID =NULL, Name = NULL){
  franchise <- "/franchise"
  req <- GET(paste0(Base_URL,franchise)) # from httr library
  data <- content(req) # from httr library

  df <- data.frame(matrix(rep(0,length(data$data[[2]])), nrow=1))
  colnames(df) <- names(data$data[[2]])
  for( i in 1:length(data$data)){
    datalist <- data$data[[i]]
    temp <- data.frame(matrix(datalist, nrow=1))
    df[i,] <- temp
  }

  NamesDf <- df %>% 
    select(id, teamCommonName) %>%
    filter(teamCommonName %in% Name)
  ID <- unique(union(ID,unlist(NamesDf$id )))
  
  
  if(is.null(ID)){
    ID <- c(1:39)
  }

  FranchiseGoalieRecords <- paste0("/franchise-goalie-records?cayenneExp=franchiseId=",ID[1])
  req <- GET(paste0(Base_URL,FranchiseGoalieRecords)) # from httr library
  data <- content(req) # from httr library
  df <- data.frame(matrix(data$data[[1]],nrow = 1))
  colnames(df) <- names(data$data[[1]])
  for(i in 2:length(ID)){
    FranchiseGoalieRecords <- paste0("/franchise-goalie-records?cayenneExp=franchiseId=",ID[2])
    req <- GET(paste0(Base_URL,FranchiseGoalieRecords)) # from httr library
    data <- content(req) # from httr library
    temp <- data.frame(matrix(data$data[[1]],nrow = 1))
    colnames(temp) <- names(df)
    df <- rbind(df,temp)
  }
  return(df)
}
```

```{r}
ID <-5
FranchiseSkatorRecords <- paste0("/franchise-skater-records?cayenneExp=franchiseId=",ID)
req <- GET(paste0(Base_URL,FranchiseSeasonRecords)) # from httr library
data <- content(req) # from httr library

```

```{r}
FranchiseAdminHistory <- paste0("franchise-detail?cayenneExp=mostRecentTeamId=",ID)



```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
